Supplemental Information: Using hydraulic geometry and the SWOT satellite to remotely sense river gas exchange velocity

### 2022-02-14 docs/supp_v8.Rmd
some ms and prepping for final validation

## Contents
This supplementary information contains 4 texts, 2 figures and 3 tables. Please consult https://github.com/craigbrinkerhoff/RSK600 for all code to build and generate results, figures, and the manuscript.

## Test S1: Estimating how many SWOT rivers are hydraulically wide
To quantify how frequently SWOT-observable rivers are also hydraulically wide river channels (section 3 of main text), we used the dataset of field-measured river hydraulics in @brinkerhoffReconcilingAtaStationAtManyStations2019. That dataset has over 500,000 discrete measurements of river width, velocity, area, and discharge that were made by the United States Geological Survey (USGS) to calibrate streamgauge rating curves. Here, we describe how this dataset was filtered down to 171,553 measurements and how we quantified what was 'hydraulically wide'.

@@

## Text S3: BIKER hyperparameterization
In this text we explain in detail how BIKER's hyperparameter values were set for each river.

We assign prior hyperparameters using SWOT data only. All priors are formalized within the model as truncated normal distributions of the log-transformed terms such that $log(X) \sim  \mathcal{N}(\mu, \sigma^2)$ for $\lambda < log(X) < \gamma$, using prior hyperparameters mean ($\mu$), standard deviation ($\sigma$), and upper ($\gamma$) and lower bounds ($\lambda$) for any parameter *X*.

@@

Finally, $k_{600_{\sigma}}$ is set to 0.30 (log-space). This corresponds to a coefficient of variation of approximately 0.30, which we took to reflect a reasonably strong agreement between the prior and the observed values. $k_{600_{\lambda}}$ was set to log(0.001) m/day. $k_{600_{\gamma}}$ was set to log(500) m/day.

## Text S4: Specifiying the $\sigma_{k_{600}}$ parameter
Here, we detail how the $\sigma_{k_{600}}$ parameter is specified in this study, as well as how it should be specified once SWOT launches.

Recall that $\sigma_{k_{600}}$ refers to the total uncertainty inherent in equation 8, i.e. stemming from both the Reynolds/form-drag dissipation model and the Manning's model for $\bar{U}$. For the purposes of this study, we are validating BIKER against equation 10 in the main text and so all uncertainties associated with the $\beta_1$ parameter are ignored and we only need to reflect the Manning's uncertainty in our specification of $\sigma_{k_{600}}^2$. Therefore, we take @hagemannBAMBayesianAMHGManning2017's estimated uncertainty from Manning's equation to infer streamflow from SWOT observations (0.25) and inflate it slightly to also account for the hydraulically-wide channel assumption and arrive at 0.30. This is the $\sigma_{k_{600}}$ used in this study.

@@

## Table S1
*Table S1: Studies that gas exchange velocity measurements come from. 'Study' refers to the paper from which these measurements were taken. Any data wrangling was done by those authors. 'Field Workers' refers to who actually made the measurements. The Raymond et al. (2012) dataset is itself a meta-analysis. Please see that paper for how those measurements were collected. We used the set of measurements ultimately published by Ulseth et al. 2019.*

>>>>

## Contents
This supplementary information contains 3 texts, 4 figures and 3 tables. Please consult https://github.com/craigbrinkerhoff/RSK600 for all code to build and generate results, figures, and the manuscript.

## Test S1: Estimating how hydraulically wide SWOT rivers are
To quantify the prevalence of hydraulically-wide, SWOT-observable rivers, we used the dataset of field-measured river hydraulics outlined in Section 2.1. That dataset has over 500,000 discrete measurements of river width, velocity, area, and discharge that were made by the United States Geological Survey (USGS) to calibrate streamgauge rating curves. Here, we describe how this dataset was filtered down to 171,553 measurements and how we quantified what was 'hydraulically wide'.

@@

## Text S3: BIKER hyperparameterization
In this text we explain in detail how BIKER's prior distributions were set for each river.

Prior distributions are defined by their hyperparameters. For BIKER, prior distriubtions are formalized as truncated normal distributions of the log-transformed terms such that $log(X) \sim  \mathcal{N}(\mu, \sigma^2)$ for $\lambda < log(X) < \gamma$, using prior hyperparameters mean ($\mu$), standard deviation ($\sigma$), and upper ($\gamma$) and lower bounds ($\lambda$) for any parameter *X*. It is important to again stress, as we do in the main text, that BIKER prior hyperparameters are described using only SWOT data to be completely globally implementanble. This is analogous to the 'empirical Bayes method' [@hoffFirstCourseBayesian2009].

@@

$k_{600_{\sigma}}$ is set to 0.30 (log-space). This corresponds to a coefficient of variation of approximately 0.30, which we took to reflect a reasonably strong agreement between the prior and the observed values. $k_{600_{\lambda}}$ was set to log(0.001) m/day. $k_{600_{\gamma}}$ was set to log(500) m/day.

Finally, recall that $\sigma_{k_{600}}$ refers to the total uncertainty inherent in equation 9, i.e. stemming from the Reynolds-extension model (equation 7) and Manning's equation for $\bar{U}$. For the purposes of this study, we are validating BIKER against equation 7 in the main text and so all uncertainties associated with the $\beta_1$ parameter are ignored and we only need to reflect the Manning's uncertainty in our specification of $\sigma_{k_{600}}^2$. Therefore, we take @hagemannBAMBayesianAMHGManning2017's estimated uncertainty from Manning's equation to infer streamflow from SWOT observations (0.25) and inflate it slightly to also account for the hydraulically-wide channel assumption and arrive at 0.30. This is the $\sigma_{k_{600}}$ used in this study.

@@

## Table S1
*Table S1: Studies that gas exchange velocity measurements come from. 'Study' refers to the paper from which these measurements were obtained. Any data wrangling was done by those authors. 'Workers' refers to who actually made the measurements. The Raymond et al. (2012) dataset is itself a meta-analysis. Please see that paper for how those measurements were collected, and see the 'Additional Studies' for the workers who actually collected the measurements.*
